# Literature review {#lit-rev}

This chapter provides an overview of the literature relevant to this thesis.
While the research conducted for this thesis did not involve any smart control, its aims are best understood by considering them through a smart control lens.
As such, this chapter begins by noting key literature relating to smart control off appliances, and any relevant findings.
Next, Section \@ref(edf) presents literature relating to electricity demand forecasting. Section \@ref(edf) begins by briefly reiterating how forecasting models are constructed, and then lists a selection of models used in electricity demand forecasting. It then notes other work whereby models were compared with one another, and the results of these comparisons.
Following this is Section \@ref(DHWM), which presents key literature regarding domestic hot water simulations. Section \@ref(DHWM) describes a selection of physical hot water models used in research simulations, and explains their interrelation with demand forecasting models.
Finally, Section \@ref(contrib) provides an explanation as to where the research in this thesis fits in relation to existing literature.
One piece of existing research in particular, that of [@Gelazanskas2015], inspired much of the research in this thesis. A number of its methodologies were replicated, and its key findings were built upon in this work. As such, [@Gelazanskas2015] is examined in some detail in Section \@ref(contrib), with differences between it and this work made clear.

## Smart control {#smart-cont-lit}

Smart control research generally involve estimating the benefits associated with scheduling or otherwise adjusting the operation of suitable appliances. This is generally done at key times in order to meet a number of objectives, such as:

* reducing grid peaks[@Roux2018; @Sepulveda2010, @Kepplinger2015]
* providing grid-beneficial (instantaneous response) services[@Kondoh2011]
* minimising costs under time-varying electricity prices[@Kepplinger2015; @Mohsenian-Rad2010; @OShaughnessy2018; @Sepulveda2010; @Cui2016; @Kapsalis2018]
* utilising low-carbon electricity[@Ramchurn2012; @Passenberg2016]
* maintaining consumer comfort under a budget constraint[@Yu2013]

Insulated hot water cylinders show high flexibility, and as such are good candidates for smart control[@Stephenson2017], but similar objectives can be obtained with other thermal devices such as heating and air conditioning systems[@OShaughnessy2018; @Yu2013]. This is due to the ability of thermal energy to be stored, with only slow dissipation, providing insulation is adequate. In addition, other devices such as electric vehicles and clothes washing machines/dryers may be used for smart control purposes[@Cui2016; @Mohsenian-Rad2010], as they also display some flexibility. 

In order to effectively schedule appliances under smart control, reasonable estimates must be made as to when they are required to be utilised. An electric vehicle, for example, must be sufficiently charged before the next drive, clothes must be washed and dry in time for their next use, and a hot water cylinder must be sufficiently hot before the next shower. Accurate predictions of these 'next' usage events provide a smart control system the ability to fine tune their optimisation algorithms to take full advantage of the demand flexibility without negatively impacting the end user.
As such, demand forecasting of appliances in some form is common in smart control literature[@Cui2016; @Roux2018; @Arghira2012; @Kepplinger2015; @Sepulveda2010; @Yu2013; @Kapsalis2018; @Passenberg2016; @Gelazanskas2016].

## Electricity demand forecasting {#edf}
<!--
Many hot water electricity demand forecasting techniques use similar methods to the simulated hot water demand component of the physical models mentioned at the beginning of this section.
The key difference is in their purpose. Rather than being run in order to simulate flow events for a physical model, they are projected into the future to obtain forecasts of demand.
When used in this manner, there is no need to be replicating physical characteristics of the hot water cylinder, thus these models focus only on understanding mathematical relationships within existing demand data. 
-->
Electricity demand models are constructed using existing electricity demand data, and, in some cases, other relevant data sources that span the same period of time. The models are 'fitted' to the existing dataset such that they best replicate the existing demand data under certain constraints. Once this is completed, these models may then be projected forward in time to where no electricity demand data exists in order to make predictions.

While much research is conducted relating to predicting grid-level[@DeFelice2013; @Jain2009], substation level[@Goude2014], or building level[@Popescu2016; @Gajowniczek2014; @Ahmad2014; @McLoughlin2013] demand, some work has also been done on predicting the demand of individual household appliances [@Arghira2012], including that of hot water cylinders[@Lomet2015; @Aki2016; @Denis2019].

When forecasting at grid level, research is generally focussed on effectively matching production and distribution[@Molina-Solana2017; @DeFelice2013; @Jain2009].
At building level, the prediction of energy consumption patterns are used when budgeting, negotiating and purchasing electricity contracts, and to detect electricity theft or fraud[@Molina-Solana2017; @Gajowniczek2014].
As mentioned in Section \@ref(smart-cont-lit), appliance or equipment level electricity demand forecasting is often utilised in smart control research. 
Additional purposes for appliance level forecasting include the detection of faults, and minimisation of operation and maintanence costs[@Molina-Solana2017].

There is currently no consensus on the 'optimal' method for forecasting electricity demand.
Conventional techniques that are commonly used are either some form of time series analysis such as ARIMA[@Lomet2015; @Arghira2012; @Popescu2016, @Denis2019], regression[@Wei2015; @McLoughlin2015; @Kaytez2015; @Dervilis2018; @Massana2015], or a combination of the two[@DeFelice2013].

There are also a number of AI methods used in electricity demand forecasting. 
These include:

* ANNs[@Dervilis2018; @Chitsaz2015]
* SVMs[@Aki2016; @Jain2009; @Zhang2016]
* ensemble bagging trees[@Wang2018]
* particle swarm optimisation[@Sepulveda2010]
* reinforcement learning[@Ruelens2018]
* cluster analysis[@McLoughlin2015]
* dynamic Gaussian processes[@Dervilis2018]

ANNS and SVMs are the two most common AI methods used for electricity demand forecasting[@Ahmad2014; @MatDaut2017; @Tardioli2015; @Amasyali2018; @Magoules2016].
ANNs attempt to mimic biological learning mechanisms by processing a set of training data through clusters of artificial neurons, each of which individually receive (numerical) inputs, and assign them certain weights. These weighted values are then outputted into other artificial neurons, which perform the same task, until eventually producing a final output.
The weights assigned by each neuron are then iteratively adjusted in order to minimise the error of the final output.
For further details regarding ANN models for forecasting electricity demand refer to [@Magoules2016].

Despite being in common use, literature suggests that ANNs tend to be outperformed in both accuracy and computational speed by support vector machines when used for electricity demand forecasting[@Ahmad2014]. 
For this reason, they were not included in this comparative work, with support vector machines instead selected as the comparative AI method.

While much electricity demand forecasting research only considers a single model, some work has been done that compares a selection of models with one another.
Research comparing AI methods showed SVMs tend to outperform ANNs in accuracy[@Ahmad2014] while having reasonable computational speed[@Ahmad2014; @Magoules2016], whereas ANNs in particular were noted as having high computational time[@Dervilis2018].
While a comparison between AI methods in [@Gajowniczek2014] showed a that an ANN provides slightly more accurate forecasts when predicting a full day ahead, the SVM was most accurate when predicting the next hour. 
Some comparative research considers both AI and conventional methods[@Dervilis2018; @Kaytez2015; @Massana2015; @Felice2011]. 
SVMs were compared against ANNs and linear regression in [@Kaytez2015; @Massana2015], and found to be the most accurate model.
SVMs were not considered in [@Dervilis2018], which instead found an AI forecasting method called dynamic Gaussian processes to be by far the most accurate when predicting grid-level electricity load.
Dynamic Gaussian processes however were noted in [@McLoughlin2013] as being unsuitable for building level demand forecasting, and thus, we may assume, also for appliance level forecasting.

Time series and AI forecasting methods were compared against a naive benchmark model in [@Felice2011] to forecast the electricity demand of office buildings. The effects of including additional independent variables (regressors) to these models was analysed.
The regressors chosen were (i) hour of day, (ii) hot water demand, (iii) luminosity, and (iv) workday/not workday.
The addition of the regressors increased accuracy drastically in the AI methods, and slightly in the time series methods.
Overall, the time series methods had the highest accuracy, although with the addition of the external regressors to the AI method this advantage was slight.

A comprehensive review of data-driven building energy consumption prediction studies[@Amasyali2018] shows SVMs consistantly outperforming ANNs in accuracy for demand forecasting, which is confirmed by another review in [@MatDaut2017]. 
Comparative research in [@Gelazanskas2015] considers a number of time series analysis methods for forecasting volumetric residential hot water demand, with ARIMA models featuring as highly regarded for this purpose.

Some demand forecasting models, particularly those using conventional forecasting methods, are reasonably easy to interpret. Once they have been fitted to the data, their parameters can be extracted to provide statistical insights into the behaviour behind the data.
This is valuable for building 'physical' domestic hot water computer models, as their parameters can be used to construct more accurate simulations of hot water demand, a necessary component of these models[@Lomet2015].

## Domestic hot water modelling {#DHWM}

Domestic hot water modelling refers to the process of building computer models that represent a residential hot water cylinder and its operation.
In the wider context of computer modelling, these are referred to as 'physical' models.
While all models are less accurate than the systems they replicate, analyses that utilise computer models have a number of crucial benefits over real-world alternatives.
Firstly, a well-constructed physical model provides the ability to make an analysis regarding alterations to the physical system that may be difficult or expensive to carry out in reality. 
Secondly, physical models provide the ability to ‘speed up’ the passing of time, and to be run ‘into the future’, two features currently unavailable when conducting real-world analyses. 
Hot water cylinder models may be utilised for a number of applications.
Some represent the physical system, with all its temperature gradients and heat and fluid flows, to a high degree of accuracy. These are suited for analyses as to the effects of making physical changes to the system, such as adjusting the position of the outlet[@Kong2016; @Yaici2013] or cylinder dimensions [@Yaici2013]. However, a high degree of physical accuracy comes with high computational cost.
The more physically simplistic models are more suitable for analyses that span over long periods of time, such as in simulations of smart control scenarios.

Generally speaking, a hot water cylinder model will consist of both a mathematical description of the physical system, and a mathematical or statistical description of the flow of hot water out of the tank, that is, hot water demand. 
A mathematical representation of the physical cylinder must, at a minimum, take into consideration the internal temperature effects of inflows of cold and outflows of hot water, heat input into the cylinder from the element, and heat loss from the walls of the cylinder.
Water temperature(s) within the cylinder are important, both for heat flow calculations, and for determining if and when the cylinder has run out of hot water.
In a cylinder of warm water at equilibrium a temperature difference between the top and bottom of the cylinder naturally arises. This can be described in terms of three distinct volumes of water. At the bottom of the cylinder, there is a volume of cooler water, with only a slight change in temperature from the bottom to the top of it. At the top, a volume of warmer water, again with only a slight vertical temperature gradient.
These are separated by a thin volume of water known as the thermocline, whereby a rapid change in temperature occurs between its top and bottom[@JackDew2018].

Some models choose to simplify this heat distribution within the tank, considering it instead to be one well-mixed, single temperature volume of water[@Heleno2015; @Parra2016; @Roux2017]. This method has the benefit of being computationally inexpensive, allowing for temporally extensive analyses to be carried out, to the detriment of some physical accuracy.
Other models simplify the heat distribution as two well-mixed bodies of water at the top and bottom, separated by a thermocline of negligible width[@JackDew2018; @Booysen2017; @Nel2018].
While requiring more computer resources than the single-volume models, the two-volume models still a provide relatively computationally inexpensive means of analysing hot water usage, while capturing some additional physical characteristics such as the rapid ‘running out’ of hot water.
A schematic of a two-volume model is provided in Figure \@ref(fig:twoVolume)
Finally, some hot water cylinder models use computational fluid dynamics (CFD) software to maintain very high amounts of physical accuracy, essentially allowing the determination of water temperature at any point within the cylinder[@Kong2016; @Yaici2013]. 

```{r twoVolume, out.width='100%', fig.align='center', fig.cap="A schematic representing a two-volume physical model of a hot water cylinder"}
knitr::include_graphics(paste0(pFolder, "schematic.pdf"))
```
Within these models there must be the ability to simulate the demand of hot water.
The model in [@Heleno2015] simply schedules three 35L, 38° hot water events at three separate times through the day.
The model in [@JackDew2018] simulates domestic hot water usage by assuming up to three major use events (representing showers, bathing etc) that are random and normally distributed about fixed times of day, with additional minor usage events (representing hand washing etc) having equal probability of occurring between specified intervals at other times.
For much of the hot water cylinder research previously mentioned however, the methodology around demand simulation is not elucidated. 

Statistical insights gained from parameters within appropriate demand forecasting models have to potential to be incorporated into physical models to simulate demand. This would improve how accurately these physical models represent reality[@Lomet2015], in turn improving accuracy of results from smart control simulations.
The research conducted for this thesis does not use any physical models. 
However, when the models that are used are compared to one another, suitability for incorporation into a physical model is considered.

## Model selection

Every forecasting model has unique strengths and weaknesses that render it more or less suitable to the aims of this thesis outlined in Section \@ref(Aims).
Based on suitability to meet these aims, some models encountered within existing literature were selected for further investigation, while others were disregarded.

This research places emphasis on variations of the ARIMA model, as existing literature suggested these would have a reasonable performance over all the comparative metrics. 
For the same reason, SVMs were chosen to represent AI methods.
ANNs were not investigated in any depth, as they appeared to required high capacity computational resources to execute in a timely manner while not offering accuracy improvements over SVMs. 
Predictions are made based only on data that would be available from any smart meter with a separately metered hot water cylinder, as is commonly available in New Zealand today.
For this reason, models that rely on the availability of external data were not utilised, although some regression was carried out using other appliances as the regressor variable.
<!--As this work is concerned mainly with making predictions only half an hour into the future, time series methods that attempt to uncover longer-term trends in averages are also not relevant.-->

## Contribution of this thesis to the literature {#contrib}

While there is some existing research that compares electricity demand forecasting models, comparisons generally only consider accuracy of predictions.
When computational costs are mentioned, they are qualitative, and given little focus[@Amasyali2018]. 
There is very little consideration within existing literature to utilising parameters from forecasting models in order to simulate demand within a physical model[@Lomet2015].

While the researchers in [@Felice2011] use hot water demand as a regressor to improve prediction of total building electricity demand, no research was found that used the electricity demand of other appliances as a regressor to to improve prediction of hot water electricity demand.

To the best of my knowledge, there is no existing literature that compares a range of models to forecast residential hot water electricity demand.
Within the existing literature, the piece of research most similar to that conducted for this thesis is that of Gelažanskas and Gamage[@Gelazanskas2015]. 
These researchers also compare a selection of forecasting methods to predict volumetric residential hot water demand.
These were fitted to volumetric hot water data collected from 95 UK households. 
Their research also used the R programming language, and used simplistic models as benchmarks by which to compare others to. 
They explored a number of time series models for prediction, including exponential smoothing, ARIMA, seasonal decomposition by local polynomial regression (STL), and combinations of these.
Predictions were made for 24 hours ahead, and models were compared against one another for accuracy.
Forecasts were also constructed for aggregated demand data, that is, the demand of all the households added together.
Accuracy of all models considered was greatly improved when aggregating data in this manner.
While my thesis draws on the key findings within this research, it differs in a number of ways, which are shown in Table \@ref(tab:gelazanskasComparison).

This chapter indicated that much groundwork has been done in existing literature as to which models would be good candidates for predicting residential hot water electricity demand.
A selection of promising models were chosen for comparison based on their performance and/or prevalence in existing literature. These are described in more detail in the methodology chapter that follows.

```{r gelazanskasComparison}
df1 <- data.frame(Characteristics = c("Data resolution", "Data type", "Forecast length", "Forecast methods", "Comparative metrics"),
                  This_thesis = c("Half hour", "Electricity demand", "Half hour", "Time series, causal, AI", "Various"), 
                  Gelazanskas = c("1 hour", "Volumetric demand", "24 hours", "Time series", "Accuracy"))

library(kableExtra)
df1 %>% 
 knitr::kable(col.names = c(" ", "This thesis", "Gelažanskas and Gamage"),
              caption = "Comparison between this thesis and the most similar piece of existing literature", 
              format = "latex") %>%
  row_spec(0, bold = TRUE) %>% 
#  column_spec(2, width = "3em") %>%
  kable_styling()
```


