`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 

<!--
If you feel it necessary to include an appendix, it goes here.
-->


# Tables and parameters {#tables}

## Seasonality

```{r frequencyAnalysis, echo=FALSE, out.width='100%'}
library(kableExtra)
load(file=paste0(dFolder, "freq_DF.Rda"))
freq_DF %>%
  knitr::kable(col.names = c("Household", "Dominance", "Period (hours)"),
               digits = 2,
               caption = "Three dominant cycle periods for hot water electricity demand of each household, as determined using frequency analysis",
               longtable = TRUE)
```

## Cross-covariance lag

```{r loadCCVdata, echo=FALSE}
load(paste0(dFolder, "ccv_max.Rda"))
maxCor <- maxCor[, c(2,1,3)]
```

```{r ccvMaxLagTable, echo=FALSE}
maxCor %>%
  knitr::kable(col.names =  c("Household", "Lag", "Cross covariance"),
        caption = "Lag time and correlation value at point of maximum crosscovariance",
        longtable = TRUE)
```

## Modelling parameters

```{R load_parameter_data, echo=FALSE}
ARIMApar <- read.csv(paste0(dFolder, "models/ARIMA/parameters.csv"))
ARIMAXpar <- read.csv(paste0(dFolder, "models/ARIMAX/parameters.csv"))
STLARIMApar <- read.csv(paste0(dFolder, "models/STLARIMA/parameters.csv"))
STLARIMAXpar <- read.csv(paste0(dFolder, "models/STLARIMAX/parameters.csv"))
```

```{r ARIMAparameters, echo=FALSE}
ARIMApar %>%
  knitr::kable(caption = "Optimal parameters for the ARIMA models")
```
Table \@ref(tab:ARIMAparameters) show the ARIMA parameters that minimise the AIC as selected by the `auto.arima` function.
As described in Chapter \@ref(methodology), $p$ refers to the order of autogression, $d$ refers to the order of differencing necessary to obtain stationarity, and $q$ is the order of moving average.
Similarly, Tables \@ref(tab:ARIMAXparameters), \@ref(tab:STLARIMAparameters) and \@ref(tab:STLARIMAXparameters) show the parameters for the ARIMAX, STL + ARIMA, and STL + ARIMAX models respectively.

```{r ARIMAXparameters, echo=FALSE}
ARIMAXpar %>%
  knitr::kable(caption = "Optimal parameters for the ARIMAX models")
```

```{r STLARIMAparameters, echo=FALSE}
STLARIMApar %>%
  knitr::kable(caption = "Optimal parameters for the STL + ARIMA models")
```

```{r STLARIMAXparameters, echo=FALSE}
STLARIMAXpar %>%
  knitr::kable(caption = "Optimal parameters for the STL + ARIMAX models")
```
